STA107 Post-Course Survey Analysis

### 2025-04-14 paper/paper.qmd
Added changes for thematic analysis

The post-course survey comprised 17 open- and close-ended questions. Among these, 11 were qualitative in nature (`Q3`, `Q6`, `Q8`, `Q10`, `Q11`, `Q12`, `Q13`, `Q14`, `Q15`, `Q16`, `Q17`), focusing on free-text responses reflecting students' thoughts on the content, instruction, and overall experience.

@@

# Rename columns
colnames(data)[1:11] <- c("Q3", "Q6", "Q8", "Q10", "Q11", 
                          "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")

@@

# Apply the function to all relevant questions
question_names <- c("Q3", "Q6", "Q8", "Q10", "Q11", 
                    "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")

@@

The sentiment analysis revealed a clear trend of positive student engagement across most open-ended responses. When asked about what aspects of the course were most helpful, students consistently expressed appreciation for the structured content delivery, use of R, and clarity in statistical explanations. Similarly, reflections on the usefulness of the normal model and applied examples elicited strong positive sentiment, highlighting the effectiveness of concept integration throughout the term.

Feedback on the clarity of teaching and accessibility of course material was more mixed. While many students praised the instructor’s effort and teaching style, others noted inconsistencies in understanding due to the pace of delivery or difficulty following complex R assignments.

The most critical responses were seen in prompts asking what aspects of the course were least helpful and what improvements could be made. Here, students frequently cited the fast pace of certain topics, confusion around R syntax, and a need for more real-world or scaffolded examples. Words like “confusing,” “fast,” and “examples” frequently appeared in these responses, pointing to the need for clearer instructional supports.

@@

From a qualitative perspective, student feedback revealed strong appreciation for several key elements of the course—especially the hands-on R activities, the integration of real-world statistical examples, and the clarity of normal model explanations. These components were described as engaging and instructive, affirming their value in helping students apply theoretical knowledge in practical settings.

>>>>

The post-course survey comprised 17 open- and close-ended questions. Among these, 8 were qualitative in nature (`Q3`, `Q6`, `Q8`, `Q10`, `Q11`, `Q12`, `Q13`, `Q14`, `Q15`, `Q16`, `Q17`), focusing on free-text responses reflecting students' thoughts on the content, instruction, and overall experience.

@@

# Rename columns
colnames(data)[1:8] <- c("Q3", "Q6", "Q8", "Q10", "Q13", "Q15", "Q16", "Q17")

@@

# Apply the function to all relevant questions
question_names <- c("Q3", "Q6", "Q8", "Q10", "Q13", "Q15", "Q16", "Q17")

@@

To gain deeper insights into student opinions, the qualitative analysis comprised two parts: thematic analysis and sentiment analysis.

### Thematic Analysis

Following Braun and Clarke’s (2006) framework, a sample of approximately 80 student responses was examined through six phases:

1. **Familiarization**: Students’ responses across 8 open-ended questions were read and annotated to gain an overall understanding.

2. **Generating Initial Codes**: Semantic-level codes were created to capture features related to course experience, R skill development, and content relevance.

3. **Searching for Themes**: Related codes were clustered into preliminary themes reflecting student attitudes and feedback.

4. **Reviewing Themes**: Themes were refined to ensure clarity and distinctiveness across the dataset.

5. **Defining and Naming Themes**: Clear definitions were established for each theme based on recurring patterns.

6. **Producing the Report**: Final themes were reported below with supporting quotes.

### Identified Themes

#### Relevance to Real Life and Careers
Many students expressed appreciation for statistics content when it connected to real-world examples or their field of study.

> "This course gave me a solid foundation in statistical concepts and taught me how to apply them using R. Overall, I gained practical skills and became more confident in working with data."

> "The course showed me how statistics applies to real-world problems, such as data analysis in my ENV course as it has alot of data."

#### R as a Learning Tool
Students had mixed responses about learning R. While some found it helpful, others found it initially intimidating.

> "After the course, I had a higher opinion of my ability to use R because I learned how to create visualizations and perform statistical analyses such as normal distribution applications and regression. Now I can use it to organize data, generate graphs, and interpret the output. The tutorial session and assignments were particularly helpful in helping me gradually build my skills."

> "I have never worked with R, but after trying to work with it I realized that it is also a powerful language."

#### Clarity and Structure of Assignments
Clearer instructions and expectations were a recurring theme.

> "The weekly assignment was a R assignment based on the module. It was sometimes hard to understand the purpose of the task."

> "After completing the weekly assignments, I have a general understanding of the statistical concepts in each module, but sometimes I still have some difficulties understanding some difficult concepts."

#### Appreciation for Tutorials and TA Support
Tutorials and TA help sessions were viewed positively, especially for R and concept reinforcement.

> "For each question on the tutorial worksheet, my TA explained the relevant concepts before attempting the problem, which helped me understand the concept better. In addition, all tutorial exercises were related to the weekly assignment, so it gave me reassurance that I was doing everything correctly before submitting my assignments."

> "The weekly tutorial was extremely useful for me. The worksheets were helpful and my TA, Dylan, did an excellent job at explaining the concepts involved and answering our questions when we were confused."

> "my ta was great at explaining concepts and let us ask many questions which would all get responses. very good by jaiditya."

#### Feedback and Reflection
Students wanted more timely or individualized feedback.

> "I always read the feedback give by the TA's since they allowed me to understand how to communicate my answers better. The statistical examples did help me recognize the usefulness and value of learning about statistics because it made me realize that statistics is everywhere."

> "Most of the feedback I recieved was in the form of checks or X. I feel like I could have gotten a little bit better feedback but most of the assignments I did were correct."

### Sentiment Analysis

The sentiment analysis provided a quantitative measure of student engagement across the survey’s open-ended responses. Using the NRC sentiment lexicon (implemented via the syuzhet package), we quantified positive, negative, and neutral sentiments for each question.

The analysis revealed:

- **Predominantly Positive Sentiment**: Most responses expressed positive sentiments, especially regarding structured content delivery, effective use of R, and clear statistical explanations.

- **Mixed Feedback on Clarity and Pace**: While many responses were positive overall, some negative sentiments were associated with challenges such as rapid topic progression and difficulties with R syntax.

- **Critical Comments**: Negative terms like “confusing,” “fast,” and “examples” frequently appeared in responses about areas needing clearer instructional support.

The bar chart above (generated in the sentiment analysis code chunk) illustrates the distribution of sentiment (positive, negative, neutral) for each of the 8 survey questions.

@@

From a qualitative perspective, student feedback revealed strong appreciation for several key elements of the course especially the hands-on R activities, the integration of real-world statistical examples, and the clarity of normal model explanations. These components were described as engaging and instructive, affirming their value in helping students apply theoretical knowledge in practical settings.

### 2025-04-14 paper/paper.tex
Added changes for thematic analysis

The post-course survey comprised 17 open- and close-ended questions.
Among these, 11 were qualitative in nature (\texttt{Q3}, \texttt{Q6},
\texttt{Q8}, \texttt{Q10}, \texttt{Q11}, \texttt{Q12}, \texttt{Q13},
\texttt{Q14}, \texttt{Q15}, \texttt{Q16}, \texttt{Q17}), focusing on
free-text responses reflecting students' thoughts on the content,
instruction, and overall experience.

@@

The sentiment analysis revealed a clear trend of positive student
engagement across most open-ended responses. When asked about what
aspects of the course were most helpful, students consistently expressed
appreciation for the structured content delivery, use of R, and clarity
in statistical explanations. Similarly, reflections on the usefulness of
the normal model and applied examples elicited strong positive
sentiment, highlighting the effectiveness of concept integration
throughout the term.

Feedback on the clarity of teaching and accessibility of course material
was more mixed. While many students praised the instructor's effort and
teaching style, others noted inconsistencies in understanding due to the
pace of delivery or difficulty following complex R assignments.

The most critical responses were seen in prompts asking what aspects of
the course were least helpful and what improvements could be made. Here,
students frequently cited the fast pace of certain topics, confusion
around R syntax, and a need for more real-world or scaffolded examples.
Words like ``confusing,'' ``fast,'' and ``examples'' frequently appeared
in these responses, pointing to the need for clearer instructional
supports.

@@

From a qualitative perspective, student feedback revealed strong
appreciation for several key elements of the course---especially the
hands-on R activities, the integration of real-world statistical
examples, and the clarity of normal model explanations. These components
were described as engaging and instructive, affirming their value in
helping students apply theoretical knowledge in practical settings.

>>>>

The post-course survey comprised 17 open- and close-ended questions.
Among these, 8 were qualitative in nature (\texttt{Q3}, \texttt{Q6},
\texttt{Q8}, \texttt{Q10}, \texttt{Q11}, \texttt{Q12}, \texttt{Q13},
\texttt{Q14}, \texttt{Q15}, \texttt{Q16}, \texttt{Q17}), focusing on
free-text responses reflecting students' thoughts on the content,
instruction, and overall experience.

@@

To gain deeper insights into student opinions, the qualitative analysis
comprised two parts: thematic analysis and sentiment analysis.

\subsubsection{Thematic Analysis}\label{thematic-analysis}

Following Braun and Clarke's (2006) framework, a sample of approximately
80 student responses was examined through six phases:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Familiarization}: Students' responses across 8 open-ended
  questions were read and annotated to gain an overall understanding.
\item
  \textbf{Generating Initial Codes}: Semantic-level codes were created
  to capture features related to course experience, R skill development,
  and content relevance.
\item
  \textbf{Searching for Themes}: Related codes were clustered into
  preliminary themes reflecting student attitudes and feedback.
\item
  \textbf{Reviewing Themes}: Themes were refined to ensure clarity and
  distinctiveness across the dataset.
\item
  \textbf{Defining and Naming Themes}: Clear definitions were
  established for each theme based on recurring patterns.
\item
  \textbf{Producing the Report}: Final themes were reported below with
  supporting quotes.
\end{enumerate}

\subsubsection{Identified Themes}\label{identified-themes}

\paragraph{Relevance to Real Life and
Careers}\label{relevance-to-real-life-and-careers}

Many students expressed appreciation for statistics content when it
connected to real-world examples or their field of study.

\begin{quote}
``This course gave me a solid foundation in statistical concepts and
taught me how to apply them using R. Overall, I gained practical skills
and became more confident in working with data.''
\end{quote}

\begin{quote}
``The course showed me how statistics applies to real-world problems,
such as data analysis in my ENV course as it has alot of data.''
\end{quote}

\paragraph{R as a Learning Tool}\label{r-as-a-learning-tool}

Students had mixed responses about learning R. While some found it
helpful, others found it initially intimidating.

\begin{quote}
``After the course, I had a higher opinion of my ability to use R
because I learned how to create visualizations and perform statistical
analyses such as normal distribution applications and regression. Now I
can use it to organize data, generate graphs, and interpret the output.
The tutorial session and assignments were particularly helpful in
helping me gradually build my skills.''
\end{quote}

\begin{quote}
``I have never worked with R, but after trying to work with it I
realized that it is also a powerful language.''
\end{quote}

\paragraph{Clarity and Structure of
Assignments}\label{clarity-and-structure-of-assignments}

Clearer instructions and expectations were a recurring theme.

\begin{quote}
``The weekly assignment was a R assignment based on the module. It was
sometimes hard to understand the purpose of the task.''
\end{quote}

\begin{quote}
``After completing the weekly assignments, I have a general
understanding of the statistical concepts in each module, but sometimes
I still have some difficulties understanding some difficult concepts.''
\end{quote}

\paragraph{Appreciation for Tutorials and TA
Support}\label{appreciation-for-tutorials-and-ta-support}

Tutorials and TA help sessions were viewed positively, especially for R
and concept reinforcement.

\begin{quote}
``For each question on the tutorial worksheet, my TA explained the
relevant concepts before attempting the problem, which helped me
understand the concept better. In addition, all tutorial exercises were
related to the weekly assignment, so it gave me reassurance that I was
doing everything correctly before submitting my assignments.''
\end{quote}

\begin{quote}
``The weekly tutorial was extremely useful for me. The worksheets were
helpful and my TA, Dylan, did an excellent job at explaining the
concepts involved and answering our questions when we were confused.''
\end{quote}

\begin{quote}
``my ta was great at explaining concepts and let us ask many questions
which would all get responses. very good by jaiditya.''
\end{quote}

\paragraph{Feedback and Reflection}\label{feedback-and-reflection}

Students wanted more timely or individualized feedback.

\begin{quote}
``I always read the feedback give by the TA's since they allowed me to
understand how to communicate my answers better. The statistical
examples did help me recognize the usefulness and value of learning
about statistics because it made me realize that statistics is
everywhere.''
\end{quote}

\begin{quote}
``Most of the feedback I recieved was in the form of checks or X. I feel
like I could have gotten a little bit better feedback but most of the
assignments I did were correct.''
\end{quote}

\subsubsection{Sentiment Analysis}\label{sentiment-analysis}

The sentiment analysis provided a quantitative measure of student
engagement across the survey's open-ended responses. Using the NRC
sentiment lexicon (implemented via the syuzhet package), we quantified
positive, negative, and neutral sentiments for each question.

The analysis revealed:

\begin{itemize}
\item
  \textbf{Predominantly Positive Sentiment}: Most responses expressed
  positive sentiments, especially regarding structured content delivery,
  effective use of R, and clear statistical explanations.
\item
  \textbf{Mixed Feedback on Clarity and Pace}: While many responses were
  positive overall, some negative sentiments were associated with
  challenges such as rapid topic progression and difficulties with R
  syntax.
\item
  \textbf{Critical Comments}: Negative terms like ``confusing,''
  ``fast,'' and ``examples'' frequently appeared in responses about
  areas needing clearer instructional support.
\end{itemize}

The bar chart above (generated in the sentiment analysis code chunk)
illustrates the distribution of sentiment (positive, negative, neutral)
for each of the 8 survey questions.

@@

From a qualitative perspective, student feedback revealed strong
appreciation for several key elements of the course especially the
hands-on R activities, the integration of real-world statistical
examples, and the clarity of normal model explanations. These components
were described as engaging and instructive, affirming their value in
helping students apply theoretical knowledge in practical settings.
